Results_paper1

### 2019-04-15 Paper1_results.Rmd
updated

# Loading libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

@@

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
library(viridis)
library(ggrepel)
library(plyr)
library(maptools)
library(sp)
library(geosphere)
library(marmap)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
library(rgeos)

## Functions
source("functions/earthdist.R")
```

@@

Now we select only the sites occuring along the west coast. The purpose of this research task is to assess whether or not the intensity of upwelling varies at va range of distances from the coastline. The data used here is obtained from different satellites and thus possess different resolutions. Here I attempt to find an overlapping time series for all of the sites along the west coast. The sites here vary between upwelling centres and further away from the upwelling centre

```{r}
site_list_sub <- site_list %>%
  filter(coast == "wc") %>%
  filter(length > 3650) 
site_list_sub <- site_list_sub[c(-2, -3, -5, -6, -7,-9, -12, -13, -14),] # Here I keep all the sites with temperature obtained up to 2017
save(site_list_sub, file = "Data_P1/site_list_sub.Rdata")

@@

In this thesis, different remotely-sensed SST datasets are used. Hence, these datasets represent different resolutions at which data was obtained.

@@

## Matching the in-situ data with the G1SST SST data
insitu_MUR <- match_func(df = MUR_SST) %>%
  drop_na()
# save(insitu_MUR, file = "Data_P1/insitu_MUR.RData")
MUR_plot <- match_plot(df = insitu_MUR)
MUR_plot <- temp_plot(df = insitu_MUR)
```

@@

## Matching the Insitu data with the G1SST SST data
insitu_G1SST <- match_func(df = G1SSTsub_SST) 
# save(insitu_G1SST, file = "Data_P1/insitu_G1SST.RData")
G1SST_plot <- match_plot(df = insitu_G1SST )
G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
```

@@

## Matching the Insitu data with the K10 SST data
insitu_K10 <- match_func(df = K10_SST) 
# save(insitu_K10, file = "Data_P1/insitu_K10.RData")
K10_plot <- match_plot(df = insitu_K10)
k10_temp_plot <- temp_plot(df = insitu_K10)
```

@@

## Here I find the nearest SST pixels

Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

@@

## Matching the Insitu data with the OISST SST data
insitu_OISST <- match_func(df = OISST_sites)
# save(insitu_OISST, file = "Data_P1/insitu_OISST.RData")
OISST_plot <- match_plot(df = insitu_OISST)
OISST_temp_plot <- temp_plot(df = insitu_OISST)
```

## Wavelets

Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2014 and 2015. Thereafter, wavelet analyses ae created per season for each individual site along the west coast of South Africa. Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.

### Load site list SACTN data
Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS).

```{r}
load("Data_P1/site_list_v4.2.RData")
load("Data_P1/SACTN_daily_v4.2.RData")
```

## Load in the rest of the datasets

@@

# Time series of one year

Examining the intensity and duration of upwelling for the years 2014 - 2015

@@

############################################################################################################################################

# Calculating a shore transect
The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km). Then create wavelet analyses (or another analyses) where I compare how the intensity and duration of upwelling vary at different distances from the coastline. So would I have a dataset with temperature at 15km then at 25km then at 35km and then at 50km and so on...?

@@

transect.pixel.isobath <- function(site, distance = 25000, isobath = 999){ # Only at a distance of 25000km
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres until the chosen isobath is reached
  pixels <- data.frame()
  deep <- 999
  while(deep > isobath){
    # Find depth at distance step
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distance))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, depth = bathy2$depth)
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
    deep <- bathy2$depth
  }
  pixels2 <- pixels[pixels$depth > -999,] # This may remove the only row of data
  if(nrow(pixels2) < 1){
    pixels2 <- data.frame(site, depth = NA)
  }else{
    pixels2 <- pixels2
  }
  return(pixels2)
}

@@

# Upwelling correlated with wind patterns
## Loading in the Wind data

@@

# Slecting the important columns for each of the datasets
wind_fix <- function(df){
wind <- df%>% 
  select(station_name, date, hour, dir, speed)
}
wind_1 <- wind_fix(df = wind_1)
wind_2 <- wind_fix(df = wind_2)
wind_3 <- wind_fix(df = wind_3)

@@

wind_data <- wind_data %>% 
  group_by(date,hour,temp_sites) %>% 
  mutate(date = as.Date(date)) 
# write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
# save(wind_data, file = "Data_P1/wind_data.RData")

@@

wind_daily <- wind_data %>% 
  mutate(date = as.Date(date))%>%
   group_by(temp_sites, date) %>%
   summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
             mean_speed = round(mean(speed),2))
save(wind_daily, file = "data/wind_daily.RData")

# Loading in the daily wind data
load("Data_P1/wind_daily.RData")
```

@@

>>>>

# Loading libraries
First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

@@

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
library(viridis)
library(ggrepel)
library(plyr)
library(maptools)
library(sp)
library(geosphere)
library(marmap)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
library(rgeos)
library(ncdf4)
library(reshape2)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)

## Functions
source("functions/earthdist.R")
source("functions/wind.rose.R")
```

@@

Now we select only the sites occuring along the west coast. The purpose of this is to assess whether or not the intensity of upwelling varies at a range of distances from the coastline. The data used here is obtained from different satellites and thus represent different resolutions. I then find an overlapping time series for each of the sites along the west coast. The length of time series for each of the sites is greater than 10 years. The sites along the west coast vary between upwelling centres and further away from the upwelling centres. As a result it is hypothesised that sites located within the upwelling centre may resemble a longer more intense upwelling event when compared to sites located further away from the upwelling centre. For this analyses a long term time series of 30years is not required and so our time series starts in 1992 - 2015.

```{r}
site_list_sub <- site_list %>%
  filter(coast == "wc") %>%
  filter(length > 3650) 
site_list_sub <- site_list_sub[c(-2, -3, -5, -6, -7,-9, -12, -13, -14),] # Here I keep all the sites with temperature obtained up to 2017
# save(site_list_sub, file = "Data_P1/site_list_sub.Rdata")

@@

# Remotely sensed SST datasets
Matching the satellite sea surface temperature data point to the in situ temperature collected data points.

@@

## Matching the in-situ data with the G1SST SST data
insitu_MUR <- match_func(df = MUR_SST) %>%
  drop_na()
# save(insitu_MUR, file = "Data_P1/insitu_MUR.RData")

MUR_plot <- match_plot(df = insitu_MUR)
MUR_plot <- temp_plot(df = insitu_MUR)
```

@@

## Matching the Insitu data with the G1SST SST data
insitu_G1SST <- match_func(df = G1SSTsub_SST) 
# save(insitu_G1SST, file = "Data_P1/insitu_G1SST.RData")

G1SST_plot <- match_plot(df = insitu_G1SST )
G1SST_temp_plot <- temp_plot(df = insitu_G1SST)
```

@@

## Matching the Insitu data with the K10 SST data
insitu_K10 <- match_func(df = K10_SST) 
# save(insitu_K10, file = "Data_P1/insitu_K10.RData")

K10_plot <- match_plot(df = insitu_K10)
k10_temp_plot <- temp_plot(df = insitu_K10)
```

@@

## The nearest SST pixels
Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites.

@@

## Matching the Insitu data with the OISST SST data
insitu_OISST <- match_func(df = OISST_sites)
# save(insitu_OISST, file = "Data_P1/insitu_OISST.RData")

OISST_plot <- match_plot(df = insitu_OISST)
OISST_temp_plot <- temp_plot(df = insitu_OISST)
```

Loading in the matched satellite temperature data

@@

## Wavelets
Wavelet anlyses showing the temperature variation between sites along the west coast for the years 2010 - 2014. Working with remotely-sensed SST and *in-situ* collected coastal seawater temperature, in the analyses bellow I aim to show how the intensity and duration of upwelling events vary between sites along the west coast of South Africa. It is important to remember that each of the datasets used within this study obtained SSTs at different resolutions.

## Load in the datasets

# Time series

Creating a time series from the year 2010 to 2014 for both the satellite and in situ collected SST data. This is done to provide an accurate comparison between datasets given that the different datasets have different starting and ending dates for which temperature was collected. E.g. Temperature for the MUR  only started in 2002

@@

#########################################################################################################################

# Calculating a shore transect

The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km). Then create wavelet analyses (or another analyses) where I compare how the intensity and duration of upwelling vary at different distances from the coastline. So would I have a dataset with temperature at 15km then at 25km then at 50km and so on...?

@@

transect.pixel.isobath <- function(site, distance = 25000, isobath = 999){ # Only at a distance of 25000km
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres 
  pixels <- data.frame()
  deep <- 999
  while(deep > isobath){
    # Find depth at distance step
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distance))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, depth = bathy2$depth)
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
    deep <- bathy2$depth
  }
  pixels2 <- pixels[pixels$depth > -999,] 
  if(nrow(pixels2) < 1){
    pixels2 <- data.frame(site, depth = NA)
  }else{
    pixels2 <- pixels2
  }
  return(pixels2)
}

@@

# Wind data

Loading the Wind data. The wind dataset was supplied by the SAWS in .text format. Wind data was provided for the sites closes to the requested site as wind data may not be available for the 6 study sites in this research.  The wind direction was collected every 2 hours and by using the circular function we calcuated the daily wind direction and mean speed. There is a numerically large gap between 360 and 2 where as in degrees its not as large. The circular mean function returns the mean direction of a vector of circular data. https://cran.r-project.org/web/packages/circular/circular.pdf (Pg 130). The circular function creates circular objects around the wind direction.

@@

# Slecting the important columns for each of the datasets
wind_fix <- function(df){
wind <- df%>% 
  select(station_name, date, hour, dir, speed) #column names
}
wind_1 <- wind_fix(df = wind_1)
wind_2 <- wind_fix(df = wind_2)
wind_3 <- wind_fix(df = wind_3)

@@

wind_data <- wind_data %>% 
  group_by(date,hour,temp_sites) %>% 
  mutate(date = as.Date(date))

# write.csv(wind_data, file = "Data_P1/wind_data.csv", row.names = T)
# save(wind_data, file = "Data_P1/wind_data.RData")

@@

wind_daily <- wind_data %>% 
  mutate(date = as.Date(date))%>%
   group_by(temp_sites, date) %>%
   summarise(dir_circ = round(mean.circular(circular(dir, units = "degrees")),2),
             mean_speed = round(mean(speed),2))

# save(wind_daily, file = "Data_P1/wind_daily.RData")
# Loading in the daily wind data
load("Data_P1/wind_daily.RData")

# Subsetting to particular sites 
selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay", "Yzerfontein", "Kommetjie")

wind_daily_sub <- wind_daily %>% 
  dplyr::rename(site = temp_sites) %>% 
  filter(site %in% selected_sites) # ***Does not detect Saldanha Bay???****

# save(wind_daily_sub, file = "Data_P1/wind_daily_sub.RData")
# Loading in the daily wind data
# load("Data_P1/wind_daily_sub.RData")
```

Matching wind data to the in situ collected data. This is done in order to plot the wind patterns for the same date at which temperature was collected. The satellite obtained SST data is also matched by date to the in situ temperature data and so the dates correspond.

```{r}
insitu_wind <- match_func(df = wind_daily_sub) %>%
  drop_na()
insitu_wind <- insitu_wind %>% 
  filter(year(date) %in% seq(2010, 2014)) # The years 2010-2014 is the years of interest and the years for which wavelets are/will be created.

# save(insitu_wind, file = "Data_P1/insitu_wind.RData")
# Loading in the daily wind data
# load("Data_P1/insitu_wind.RData")
```

# Plotting the wind data - Waverose diagrams

```{r}
wave_daily_renamed <- insitu_wind %>% 
  dplyr::rename(spd = mean_speed) %>%
  dplyr::rename(dir = dir_circ)

p.wr2 <- plot.windrose(data = wave_daily_renamed,
              spd = "spd",
              dir = "dir")
p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = "") +
  theme(strip.text.x = element_text(size = 25))
p.wr3
```

## Chlorophyll-a data

Extract the chlorophyll a netCDF files and convert it to CSV. The chlorophylla data were obtained from MODIS Aqua. The aim of working with these chlorophyll a data is to examine how the chlorophyll a concentration varies with upwelling. This function convert chlorophyll netCDF datasets to a single CSV dataset.

```{r}
# MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
# MODIS_chlor.csv.dir <- "/home/amieroh/Documents/Data/Datasets"

region <- "BC" # Benguela Current
coords <- bbox[, region]
coords <- c(-35, -20, 10, 20) # this is the BC

ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
nc.init <- nc_open(ncList[1])
LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
nc_close(nc.init)

ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
nc <- nc_open(nc.file)
  instrument <- ncatt_get(nc, 0, "instrument")$value
  platform <- ncatt_get(nc, 0, "platform")$value
  product_name <- ncatt_get(nc, 0, "product_name")$value
  fNameStem <- substr(product_name, 17, 38)
  timeStamp <- substr(product_name, 2, 8)
  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
  chl <- round(ncvar_get(nc,
                   varid = "chlor_a",
                   start = c(LonIdx[1], LatIdx[1]),
                   count = c(length(LonIdx), length(LatIdx))),
               3)
  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
                        lat =  nc$dim$lat$vals[LatIdx])
  nc_close(nc)
  chl <-
    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
    mutate(t = date) %>%
    na.omit()
  fwrite(chl,
         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
         append = TRUE, col.names = FALSE)
  rm(chl)
}

llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)

# MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
#             col.names = c("lon", "lat", "chloro", "date"))

# save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
```

## Setting up the chlorophylla dataset

# Loading the data
load("Data_P1/MODIS_Chloro.RData") # Extracted chlorophyll data
load("Data_P1/site_list_sub.Rdata") # Site list of all the sites along the west coast

# Visualisation
chloro_plot <- MODIS_Chloro %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = chloro))

unique_pixel <- MODIS_Chloro %>% 
  select(lon, lat) %>% 
  unique()

match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(site_list_sub[,5:6]), k = 1)

# Matching the chlorophylla-a lats and lons to the study sites
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = site_list_sub$site)

chloro_match <- MODIS_Chloro %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)

chloro_match %>%
  filter(date == max(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = chloro))

chloro_sites <- chloro_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
  dplyr::mutate(date = as.Date(date))
# save(chloro_sites, file = "Data_P1/chloro_sites.RData")

# Visualisation
Chloro_plot_complete <- ggplot(chloro_sites, aes(x = date, y = chloro)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
Chloro_plot_complete

# Study period from 2010 - 2014
chloro_data <- chloro_sites %>% 
  filter(year(date) %in% seq(2010, 2014))
# save(chloro_data, file = "Data_P1/chloro_data.RData")

# Plotting only years 2014-2015
Chloro_plot_filtered <- ggplot(chloro_data, aes(x = date, y = chloro)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
Chloro_plot_filtered

# Creating daily temperatures
load("Data_P1/chloro_data.RData") # Loading in the chlorophyll-a data

daily_chloro_data <- chloro_data %>%
  dplyr::mutate(date = as.Date(date)) %>%
  dplyr::group_by(combi, lon, lat, date) %>%
  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
  dplyr::group_by(combi, lon, lat, date) %>%
  tidyr::nest() %>%
  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
                     climatologyPeriod = c("2002-07-13", "2018-12-04"),
                     maxPadLength = 8)) %>% # Adjust maxPadLength = X for longer interpolation
  dplyr::select(-data) %>%
  tidyr::unnest() %>%
  select(-doy, -seas, -thresh)
```

@@

Things to do

- Plot seasonal detrend
- chlorophyll a data
- fix windrose plot
- Stastical analyses
